|

### 2024-10-27 output/seminar-slideshow.qmd
Complete "Item Factor Analysis" slide

---
title:     |
  Parámetros multidimensionales en Teoría de Respuesta al Ítem:
  ¡Algo estamos haciendo mal!
author:    |
  <table>
    <tr>
      <td>Daniel Morillo, Ph.D.</td>
    </tr>
    <tr>
      <td>
        <center>
          [![](../www/slideshow-assets/github-logo.png){height="50"}](https://github.com/DaniMori/)&emsp;
          [![](../www/slideshow-assets/orcid.png){height="50"}](https://orcid.org/0000-0003-3021-3878)
        </center>
      </td>
    </tr>
  </table>
institute: |
  <table>
    <tr>
      <td>
        <center>
          [![](../www/slideshow-assets/logos_fac_psicologia.svg){height="120"}](https://www.uned.es/universidad/facultades/psicologia.html)
        </center>
      </td>
    </tr>
  </table>
bibliography:  ../www/Multidimensional-parameters-MCLM.bib
csl:           ../www/apa-old-doi-prefix.csl
date:          "2024-10-21"
date-meta:     "2024-11-27"
date-format:   long
editor:
  mode:        source
  markdown: 
    wrap:      sentence
    canonical: true
lang:          es
knitr:
  # opts_knit: # Does not work, see line 88
  #   root_dir:  here::here()
  opts_chunk: 
    results:   asis
    cache:     true
    autodep:   true
format:
  revealjs:
    auto-stretch:            true
    code-annotations:        hover
    df-print:                paged
    fig-cap-location:        bottom
    fig-format:              svg
    fig-asp:                 1
    fig-height:              6
    fig-width:               6
    incremental:             false
    keep-md:                 true
    link-external-newwindow: true
    margin:                  0
    self-contained:          true
    slide-number:            false
    theme:                   ../www/slideshow-assets/extra-styles.scss
    transition:              none
    view-distance:           3
    template-partials:
      - ../www/slideshow-assets/title-slide.html
editor_options: 
  chunk_output_type: console
---

@@

:::::: columns
:::: column
$$
  i_k^* = \tau_k + \mathbf{\lambda}_k^T \mathbf{F} + U_k
$$

@@

::: fragment
$$
  I_k = \left\{ \begin{matrix}
      1 &si &i_k^* \geq 0\\
      0 &si &i_k^* <    0
    \end{matrix} \right.
$$
:::
::::

::: column
```{r ifa-curve}
#| fig-height: 9

@@

latent_response <- tibble(
  point    = ERROR_POINTS |> rep(100000),
  response = rnorm(
    length(point),
    mean = ITEM_INTERCEPT + ITEM_LOADING * point,
    sd   = sqrt(item_unique_var)
  )
)

# Plot: Upper Panel - Linear regression with Gaussian error distributions
latent_response_plot <- latent_response |>
  ggplot(aes(x = response, y = factor(point), fill = ..x.. > 0))           +
  geom_density_ridges_gradient(
    scale          = .2,
    rel_min_height = 0.001,
    color          = PALETTE_BLUE["normal"]
  )                                                                        +
  labs(x = "i~k~^*^", y = NULL)                                            +
  scale_y_discrete(labels = ERROR_POINTS)                                  +
  scale_x_continuous(breaks = 0)                                           +
  scale_fill_manual(
    values = c(`FALSE` = "white", `TRUE` = unname(PALETTE_BLUE["white"]))) +
  theme(
    legend.position = "none",
    axis.title.y = element_markdown(angle = 90),
    panel.grid.major.y = element_line(linetype = "17")
  )                                                                        +
  coord_flip()

# Plot: Lower Panel - Probit curve
probit_plot <- probit_curve_data |>
  ggplot(mapping = aes(x = score, y = prob))                          +
  geom_line(color = PALETTE_BLUE["normal"], linewidth = LINE_WIDTH)   +
  labs(x = "i~k~^*^", y = "P(I~k = 1)")                               +
  scale_y_continuous(breaks = 0:1, minor_breaks = NULL, limits = 0:1) +
  scale_x_continuous(minor_breaks = ERROR_POINTS, breaks = -1:1 * 3)  +
  theme(
    axis.title   = element_markdown(),
    axis.title.y = element_markdown(angle = 90),
  )

## TODO: Correct error break points
latent_response_plot / probit_plot
```
:::
::::::

>>>>

---
title:     |
  Parámetros multidimensionales en Teoría de Respuesta al Ítem:
  ¡Algo estamos haciendo mal!
author:    |
  <table>
    <tr>
      <td>Daniel Morillo, Ph.D.</td>
    </tr>
    <tr>
      <td>
        <center>
          [![](../www/slideshow-assets/github-logo.png){height="50"}](https://github.com/DaniMori/)&emsp;
          [![](../www/slideshow-assets/orcid.png){height="50"}](https://orcid.org/0000-0003-3021-3878)
        </center>
      </td>
    </tr>
  </table>
institute: |
  <table>
    <tr>
      <td>
        <center>
          [![](../www/slideshow-assets/logos_fac_psicologia.svg){height="120"}](https://www.uned.es/universidad/facultades/psicologia.html)
        </center>
      </td>
    </tr>
  </table>
bibliography:  ../www/Multidimensional-parameters-MCLM.bib
csl:           ../www/apa-old-doi-prefix.csl
date:          "2024-10-21"
date-meta:     "2024-11-27"
date-format:   long
editor:
  mode:        source
  markdown: 
    wrap:      sentence
    canonical: true
lang:          es
knitr:
  # opts_knit: # Does not work, see line 88
  #   root_dir:  here::here()
  opts_chunk: 
    results:   asis
    cache:     true
    autodep:   true
format:
  revealjs:
    auto-stretch:            true
    code-annotations:        hover
    df-print:                paged
    fig-cap-location:        bottom
    fig-format:              svg
    fig-asp:                 1
    fig-width:               6
    incremental:             false
    keep-md:                 true
    link-external-newwindow: true
    margin:                  0
    self-contained:          true
    slide-number:            false
    theme:                   ../www/slideshow-assets/extra-styles.scss
    transition:              none
    view-distance:           3
    template-partials:
      - ../www/slideshow-assets/title-slide.html
editor_options: 
  chunk_output_type: console
---

@@

:::::: columns
:::: column
<br>

$$
  i_k^* = \tau_k + \mathbf{\lambda}_k^T \mathbf{F} + U_k
$$

@@

::: {.fragment fragment-index="1"}
$$
  I_k = \left\{ \begin{matrix}
      1 &si &i_k^* \geq 0\\
      0 &si &i_k^* <    0
    \end{matrix} \right.
$$
:::
::::

::: {.column .fragment fragment-index="1"}
```{r ifa-curve}
#| fig-asp: 1.4
#| cache: false

@@

latent_response <- tibble(
  point    = ERROR_POINTS |> rep(500000),
  response = rnorm(
    length(point),
    mean = ITEM_INTERCEPT + ITEM_LOADING * point,
    sd   = sqrt(item_unique_var)
  )
)

# Plot: Upper Panel - Linear regression with Gaussian error distributions
latent_response_plot <- latent_response |>
  ggplot(aes(x = response, y = point, fill = ..x.. > 0, group = point)) +
  geom_density_ridges_gradient(
    scale          =  .2,
    rel_min_height = 0.001,
    color          = PALETTE_BLUE["normal"]
  )                                                                     +
  labs(x = "i~k~^*^", y = NULL)                                         +
  scale_y_continuous(
    breaks       = ERROR_POINTS,
    minor_breaks = NULL,
    labels       = NULL,
    limits       = range(ERROR_POINTS) * 1.12,
    expand       = expansion()
    )           +
  scale_x_continuous(breaks = 0)                                        +
  scale_fill_manual(
    values = c(
      `FALSE` = unname(PALETTE_BLUE["white"]),
      `TRUE`  = unname(PALETTE_BLUE["normal"])
    )
  ) +
  theme(
    axis.ticks.x = element_blank(),
    axis.title.y = element_markdown(angle = 90),
    legend.position = "none",
    panel.grid.major = element_line(linetype = "17")
  )                                                                     +
  coord_flip()

# Plot: Lower Panel - Probit curve
probit_plot <- probit_curve_data |>
  ggplot(mapping = aes(x = score, y = prob))                        +
  geom_line(color = PALETTE_BLUE["normal"], linewidth = LINE_WIDTH) +
  labs(x = "E(*i~k~^*^*)", y = "P(I~k~ = 1)")                       +
  scale_y_continuous(
    breaks       = 0:1,
    minor_breaks = NULL,
    limits       = 0:1,
    expand       = expansion(mult = c(0, .05))
  )  +
  scale_x_continuous(
    breaks       = ERROR_POINTS,
    minor_breaks = NULL,
    limits = range(ERROR_POINTS) * 1.12,
    expand = expansion()
  ) +
  theme(
    axis.title       = element_markdown(),
    axis.title.y     = element_markdown(angle = 90),
    panel.grid.major = element_line(linetype = "17"),
  )

## TODO: Correct error break points and fig height
latent_response_plot / probit_plot
```
:::
::::::
